Fluent statistical computing interfaces for biological data analysis

### 2020-07-26 Rmd/06-conclusion.Rmd
resolve Matt comments

@@

The **plyranges** package develops a suite of verbs for interacting
with genomic data as a *GRanges* object. Since its release on Bioconductor,
it  has been relatively  successful: it has been downloaded `r plyranges_stats$no_downloads` times from  `r plyranges_stats$no_ip` unique IP addresses. I have also had the privilege of teaching workshops on **plyranges**
at Bioconductor conferences which also led to the development of the 
**fluentGenomics** workflow package, outlined in chapter \@ref(ch:fluentGenomics). A broader impact of the work, has been the discussions around the
concepts of fluent interfaces and tidy data within the Bioconductor community,
which has led to several developments currently in place that are exploring
different approaches for fluent interfaces for other types of omics data.
The **plyranges** package is available to download from 
https://bioconductor.org/packages/release/bioc/html/plyranges.html and
the **fluentGenomics** workflow is available to download from 
https://bioconductor.org/packages/release/workflows/html/fluentGenomics.html.

@@

Finally, the **liminal** software aims to provide a more holistic approach to
analysis tasks requiring the use of dimensionality reduction algorithms. We
showed how to incorporate interactive graphics and tours to identify problems
with embeddings. Based on the case studies provided I believe that the methods
used in **liminal** could be broadly applicable many high dimensional datasets.
The **liminal** package is available to download from https://github.com/sa-lee/liminal.

@@

A limitation of the grammar as we have implemented it in 
**plyranges** is lack of scalability and computational speed for data sets
that do not fit in memory. We attempted several techniques for performing 
delayed operations over range-based data, however a more general approach
that allows for data stored on the cloud or in scientific data formats
like HDF5 that leverage existing Bioconductor frameworks would be useful.
We showed in chapters \@ref(ch:fluentGenomics) ad \@ref(ch:intron) that an a
analyst is able to do some very complex data transformations and re-sampling 
procedures via casting results into *GRanges* object. However, it is unclear 
whether the semantics of our grammar can be extended to data that can not be 
easily reshaped into long form tidy representations. Moreover, further work
is required to explore the design space of grammars for data 
transformations and grammars for graphics when the data large, multifaceted
and non-rectangular.

We showed in chapter \@ref(ch:tsne) that tours provide a global overview
that can be used as tool for exploring model fits. An issue that arises is
how to scale the tour as the number of observations increases. There are
latencies in sending data from the back end to the visualisation client
that causes lag during animation. One could also question whether point
based displays are appropriate in this case, and it would be worth exploring 
the usability of animations based on binning the projections. Moreover, 
when the number of observations are large, the points in the projections are 
concentrated in the centre of the tour display obscuring interesting aspects
of the data. This is mitigated via having
the ability to zoom, but further research into transforming the projections
to avoid crowding would be valuable. An added complexity to changes in visual
displays are thinking about the design of user interactions, several promising
avenues based on section tours should be explored [@Laa2020-wr; @Laa2020-fy].

>>>>

nice <- function(x) prettyNum(x, big.mark = ",", scientific = FALSE)

@@

The **plyranges** package develops a suite of verbs for interacting
with genomic data as a *GRanges* object. Since its release on Bioconductor,
it  has been relatively  successful: it has been downloaded `r nice(plyranges_stats$no_downloads)` times from  `r nice(plyranges_stats$no_ip)` unique IP addresses. I have also had the privilege of teaching workshops on **plyranges**
at Bioconductor conferences which also led to the development of the 
**fluentGenomics** workflow package, outlined in chapter \@ref(ch:fluentGenomics). A broader impact of the work, has been the discussions around the
concepts of fluent interfaces and tidy data within the Bioconductor community,
which has led to several developments currently in place that are exploring
different approaches for fluent interfaces for other types of omics data.
The **plyranges** package is available to download from 
https://bioconductor.org/packages/release/bioc/html/plyranges.html and
the **fluentGenomics** workflow is available to download from 
https://bioconductor.org/packages/release/workflows/html/fluentGenomics.html.

@@

Finally, the **liminal** software aims to provide a more holistic approach to
analysis tasks requiring the use of dimensionality reduction algorithms. We
showed how to incorporate interactive graphics and tours to identify problems
with embeddings. Based on the case studies provided I believe that the methods
used in **liminal** could be broadly applicable to many high dimensional datasets
and NLDR methods.
The **liminal** package is available to download from https://github.com/sa-lee/liminal.

@@

A limitation of the grammar as we have implemented it in 
**plyranges** is lack of scalability and computational speed for data sets
that do not fit in memory. We attempted several techniques for performing 
delayed operations over range-based data, however a more general approach
that allows for data stored on the cloud or in scientific data formats
like HDF5 that leverage existing Bioconductor frameworks would be useful.
We showed in chapters \@ref(ch:fluentGenomics) ad \@ref(ch:intron) that an
analyst is able to do some very complex data transformations and re-sampling 
procedures via casting results into _GRanges_ object. However, it is unclear 
whether the semantics of our grammar can be extended to data that can not be 
easily reshaped into long form tidy representations. Moreover, further work
is required to explore the design space of grammars for data 
transformations and grammars for graphics when the data large, multifaceted
and non-rectangular.

We showed in chapter \@ref(ch:tsne) that tours provide a global overview
that can be used as tool for exploring model fits. An issue that arises is
how to scale the tour as the number of observations increases. There are
latencies in sending data from the back end to the visualisation client
that causes lag during animation. One could also question whether point
based displays are appropriate in this case, and it would be worth exploring 
the usability of animations based on binning the projections. Moreover, 
when the number of observations are large, the points in the projections are 
concentrated in the centre of the tour display obscuring interesting aspects
of the data. This is mitigated via having
the ability to zoom, but further research into transforming the projections
to avoid crowding would be valuable. An added complexity to changes in visual
displays are thinking about the design of user interactions, and several promising
avenues based on section tours could be explored [@Laa2020-wr; @Laa2020-fy].
